RDF Validation of Metadata \\ on Highly-Complex Person-Level Data

### 2015-02-23 paper/paper7/paper7-appendix.tex
added constraints

	Iteration over all concepts in the vocabulary and returning that don't have associated resources using (subproperties of) \emph{skos:semanticRelation}. 

@@

	Implementation: Creation of an undirected graph that includes all non-orphan concepts as nodes and all semantic relations as edges. Tarjan's algorithm then finds and returns all weakly connected components.

@@

	Implementation: Construction of a graph having all concepts as nodes and the set of edges being \emph{skos:broader} relations. 

@@

	Implementation:

@@

@@

	Implementation: Identification of all concept pairs that are related by \emph{skos:broaderTransitive} or \emph{skos:narrowerTransitive} properties but not by their \emph{skos:broader} and \emph{skos:narrower} subproperties. 

@@

	Severity level: SHOULD

@@

	Severity level: ERROR

@@

\begin{itemize}
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-01\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Undocumented Concepts}:}}
	Undocumented Concepts:
  The SKOS standard defines a number of properties useful for documenting the meaning of the concepts in a thesaurus also in a human-readable form. Intense use of these properties leads to a well-documented thesaurus which should also improve its quality.  
	Iteration over all concepts in the vocabulary and find those not using one of \emph{skos:note}, \emph{skos:changeNote}, \emph{skos:definition}, \emph{skos:editorialNote}, \emph{skos:example}, \emph{skos:historyNote}, or \emph{skos:scopeNote}.
  Severity level: INFO
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-02\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Overlapping Labels}:}}
	Overlapping Labels:
	This is a generalization of a recommendation in the SKOS primer, that “no two concepts have the same preferred lexical label in a given language when they belong to the same concept scheme”. This could indicate missing disambiguation information and thus lead to problems in autocompletion application. 
	Severity level: INFO
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-03\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Missing Labels}:}}
	Missing Labels:
	To make the vocabulary more convenient for humans to use, instances of SKOS classes (Concept, ConceptScheme, Collection) should be labeled using e.g., \emph{skos:prefLabel}, \emph{altLabel}, \emph{rdfs:label}, \emph{dc:title}. 
	Severity level: INFO
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-04\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Unprintable Characters in Labels}:}}
	Unprintable Characters in Labels:
	\emph{pref/alt/hiddenlabels} contain characters that are not alphanumeric characters or blanks.
	Severity level: INFO
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-05\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Empty Labels}:}}
	Empty Labels:
	Labels also need to contain textual information to be useful, thus we find all SKOS labels with length 0 (after removing whitespaces). 
	Severity level: INFO
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-06\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Ambiguous Notation References}:}}
	Ambiguous Notation References:
	Concepts within the same concept scheme should not have identlical \emph{skos:notation} literals. 
	Severity level: INFO
\end{itemize}

>>>>

		\item Implementation: Iteration over all concepts in the vocabulary and returning that don't have associated resources using (subproperties of) \emph{skos:semanticRelation}. 
		\item Severity level: INFO

@@

		\item Implementation: Creation of an undirected graph that includes all non-orphan concepts as nodes and all semantic relations as edges. Tarjan's algorithm then finds and returns all weakly connected components.
		\item Severity level: INFO

@@

		\item Implementation: Construction of a graph having all concepts as nodes and the set of edges being \emph{skos:broader} relations. 
		\item Severity level: INFO

@@

		\item Implementation:

@@

	  \item Severity level: INFO

@@

		\item Implementation: Identification of all concept pairs that are related by \emph{skos:broaderTransitive} or \emph{skos:narrowerTransitive} properties but not by their \emph{skos:broader} and \emph{skos:narrower} subproperties. 
		\item Severity level: INFO

@@

	  \item Severity level: SHOULD

@@

	  \item Severity level: ERROR

@@

\begin{itemize}
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-01\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Undocumented Concepts}:}}
	Undocumented Concepts:
  The SKOS standard defines a number of properties useful for documenting the meaning of the concepts in a thesaurus also in a human-readable form. Intense use of these properties leads to a well-documented thesaurus which should also improve its quality.  
	Iteration over all concepts in the vocabulary and find those not using one of \emph{skos:note}, \emph{skos:changeNote}, \emph{skos:definition}, \emph{skos:editorialNote}, \emph{skos:example}, \emph{skos:historyNote}, or \emph{skos:scopeNote}.
	\begin{itemize}
		\item Severity level: INFO
	\end{itemize}
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-02\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Overlapping Labels}:}}
	Overlapping Labels:
	This is a generalization of a recommendation in the SKOS primer, that “no two concepts have the same preferred lexical label in a given language when they belong to the same concept scheme”. This could indicate missing disambiguation information and thus lead to problems in autocompletion application. 
	\begin{itemize}
		\item Severity level: INFO
	\end{itemize}
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-03\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Missing Labels}:}}
	Missing Labels:
	To make the vocabulary more convenient for humans to use, instances of SKOS classes (Concept, ConceptScheme, Collection) should be labeled using e.g., \emph{skos:prefLabel}, \emph{altLabel}, \emph{rdfs:label}, \emph{dc:title}. 
	\begin{itemize}
		\item Severity level: INFO
	\end{itemize}
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-04\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Unprintable Characters in Labels}:}}
	Unprintable Characters in Labels:
	\emph{pref/alt/hiddenlabels} contain characters that are not alphanumeric characters or blanks.
	\begin{itemize}
		\item Severity level: INFO
	\end{itemize}
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-05\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Empty Labels}:}}
	Empty Labels:
	Labels also need to contain textual information to be useful, thus we find all SKOS labels with length 0 (after removing whitespaces). 
	\begin{itemize}
		\item Severity level: INFO
	\end{itemize}
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-06\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Ambiguous Notation References}:}}
	Ambiguous Notation References:
	Concepts within the same concept scheme should not have identlical \emph{skos:notation} literals. 
	\begin{itemize}
		\item Severity level: INFO
	\end{itemize}
\end{itemize}

\subsection{\underline{Vocabulary}}
Vocabularies should not invent any new terms or use deprecated elements.

\begin{itemize}
	\item \textbf{\em DATA-CUBE-C-VOCABULARY-01} 
	\begin{itemize}
		\item Severity level: FATALERROR
	\end{itemize}
\end{itemize}

\begin{itemize}
	\item \textbf{\em DCAT-C-VOCABULARY-01} 
	\begin{itemize}
		\item Severity level: FATALERROR
	\end{itemize}
\end{itemize}

\begin{itemize}
	\item \textbf{\em DISCO-C-VOCABULARY-01} 
	\begin{itemize}
		\item Severity level: FATALERROR
	\end{itemize}
\end{itemize}

\begin{itemize}
	\item \textbf{\em PHDD-C-VOCABULARY-01} 
	\begin{itemize}
		\item Severity level: FATALERROR
	\end{itemize}
\end{itemize}

\begin{itemize}
	\item \textbf{{\em SKOS-C-VOCABULARY-01\footnote{Corresponds to qSKOS Quality Issues - Linked Data Specific Issues - Undefined SKOS Resources}:}}
	Undefined SKOS Resources:
	The vocabulary should not invent any new terms within the SKOS namespace or use deprecated SKOS elements. 
	\begin{itemize}
		\item Severity level: FATALERROR
	\end{itemize}
\end{itemize}

\begin{itemize}
	\item \textbf{\em XKOS-C-VOCABULARY-01} 
	\begin{itemize}
		\item Severity level: FATALERROR
	\end{itemize}
\end{itemize}

### 2015-02-23 paper/paper7/paper7.tex
added constraints

- SKOS validation of hierarchies of theoretical concepts

@@

\emph{Property Domains}\footnote{\emph{R-25-OBJECT-PROPERTY-DOMAIN}, \emph{R-26-DATA-PROPERTY-DOMAIN}} and \emph{Ranges}\footnote{\emph{R-28-OBJECT-PROPERTY-RANGE}, \emph{R-35-DATA-PROPERTY-RANGE}}restrict domains and ranges of properties.
Only \emph{skos:ConceptSchemes}, e.g., can have \emph{skos:hasTopConcept} relationships (\ms{$\exists$ hasTopConcept.$\top$ $\sqsubseteq$ ConceptScheme}) and
\emph{xkos:belongsTo} relationships can only point to instances of the class \emph{skos:Concept} (\ms{$\top$ $\sqsubseteq$ $\forall$ belongsTo.Concept}).
A \emph{universal quantification}\footnote{\emph{R-91-UNIVERSAL-QUANTIFICATION-ON-PROPERTIES}} contains all those individuals that are connected by a property only to individuals/literals that are instances of particular classes  or data ranges.
Only \emph{dcat:Catalogs}, e.g., can have \emph{dcat:dataset} relationships to \emph{dcat:Datasets} (\ms{Catalog $\sqsubseteq$ $\forall$ dataset.Dataset}).
An \emph{existential quantification}\footnote{\emph{R-86-EXISTENTIAL-QUANTIFICATION-ON-PROPERTIES}} contains all those individuals that are connected by a property to individuals/literals that are instances of given classes or data ranges.
Every \emph{qb:SliceKey}, e.g., must be associated with (\emph{qb:sliceKey}) a \emph{qb:DataStructureDefinition} (\ms{SliceKey $\sqsubseteq$ $\exists$ sliceKey$^{-}$.DataStructureDefinition}).

>>>>

SKOS is based on RDF, which is a graph-based data model. Therefore, we can concentrate on the vocabulary's graph-based structure for assessing the quality of SKOS vocabularies and apply graph- and network-analysis techniques (\emph{structure}) like 
(1) a vocabulary should provide entry points (top concepts) to the data to provide efficient access and guidance for human users and
(2) concepts, internal to the tree, should not be indicated as top concepts, and
(3) a vocabulary should not contain many orphan concepts 
(An orphan concept is a concept without any associative or hierarchical relations lacking valuable context information. A controlled vocabulary that contains many orphan concepts is less usable for search and retrieval use cases, as, e.g., no hierarchical query expansion can be performed on search terms to find documents with more general content.) \cite{MaderHaslhoferIsaac2012}. 
\tb{add disco example}
- SKOS validation of hierarchies of theoretical concepts

@@

Vocabularies should not invent any new terms or use deprecated elements.
The \emph{vocabulary} constraint should be the first one to be checked for each reused vocabulary.
\emph{Property Domains}\footnote{\emph{R-25-OBJECT-PROPERTY-DOMAIN}, \emph{R-26-DATA-PROPERTY-DOMAIN}} and \emph{Ranges}\footnote{\emph{R-28-OBJECT-PROPERTY-RANGE}, \emph{R-35-DATA-PROPERTY-RANGE}}restrict domains and ranges of properties.
Only \emph{skos:ConceptSchemes}, e.g., can have \emph{skos:hasTopConcept} relationships (\ms{$\exists$ hasTopConcept.$\top$ $\sqsubseteq$ ConceptScheme}) and
\emph{xkos:belongsTo} relationships can only point to instances of the class \emph{skos:Concept} (\ms{$\top$ $\sqsubseteq$ $\forall$ belongsTo.Concept}).
A \emph{universal quantification}\footnote{\emph{R-91-UNIVERSAL-QUANTIFICATION-ON-PROPERTIES}} contains all those individuals that are connected by a property only to individuals/literals that are instances of particular classes  or data ranges.
Only \emph{dcat:Catalogs}, e.g., can have \emph{dcat:dataset} relationships to \emph{dcat:Datasets} (\ms{Catalog $\sqsubseteq$ $\forall$ dataset.Dataset}).
An \emph{existential quantification}\footnote{\emph{R-86-EXISTENTIAL-QUANTIFICATION-ON-PROPERTIES}} contains all those individuals that are connected by a property to individuals/literals that are instances of given classes or data ranges.
Every \emph{qb:SliceKey}, e.g., must be associated with (\emph{qb:sliceKey}) a \emph{qb:DataStructureDefinition} (\ms{SliceKey $\sqsubseteq$ $\exists$ sliceKey$^{-}$.DataStructureDefinition}).
